Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-05-26 review/answer_to_review_2.md

Two major take aways: (1) Their mobility estimator could be used for a better, region independent statistic for the general degree of interaction. (2) I could as well use a moving window instead of 200 years time slices (Figure 6).

I added some sentences in the conclusion to give an outlook on future work.

> I also still have some issues with how sample size is effectively being ignored once proportions are computed, and I think this should be emphasised more strongly as I expect not everyone would be reading beyond the conclusion section.

I changed the wording of the conclusion to make this very clear.

@@

> Page 5. I still think a permutation test would be more robust in this case. One could compare the observed correlation between two pairs of proportion time series against a simulation set where each iteration is the measured correlation between the proportion time-series of two artificial sets generated by randomly permuting the location label of each radiocarbon date. Pairs with a smaller number of dates will generate a wider range of correlation values and will be harder to test. One could adopt a more sophisticated solution by permuting groups of dates that occurred together in order to handle sample interdependence. Might be worth considering doing this for a follow-up study.

Well noted for the follow-up study. I also mention this approach now in the conclusion.

> Page 6, Column 2, Lines 46-49. ‘If one believes the sample, then a sudden, radical change from burial in flat graves to hill graves took place in Southern Germany at the beginning of the Middle Bronze Age.’  The sample size here is 8 dates. I appreciate the cautionary ‘if one believes the sample’ but this is a bit of a stretch to say the least, and readers that are numerically less literate might be misled. The same applies to other contexts with extremely small sample sizes and I think adding confidence intervals to those proportions would help.

I modified the small, introductory paragraph of this section to make it clear, that this is a thought experiment based on a very small amount of data. I also set the narrative in italics.

@@

> Table 3 & Figure 6. The figure nicely summarises the trend detailed on Table 3, but does not distinguish instances where the correlation were identified as signficant vs those that were not. Indeed in no cases the analysis shows a signficant correlation under the Partial Mantel test, whilst only few chronological blocks are showing signficant correlations. I suggest highlighting this by perhaps changing the the point symbology. It is also worth noting that if anything there is a bias towards type I error given: a) that we are dealing with multiple testing; and b) these analyses do not take into account sample sizes which are often very small. P-values are not the core issue here, but rather the fact that some way to at least visually convey this uncertainty is much needed.

To address this issue I modified Figure 6. I added...

1. ... an additional graph Figure 6 A that shows the median (with an lower to upper quartile ribbon) of the graves per year and region.

2. ... an indication of the mean of this medians for each 200 years time window in Figure 6 B. The square markers show on how much data the correlation observations are based.

3. ... an indication of the p-values in Figure 6 B to show which correlations are significant.

I also modified the description of this figure in the Materials and Methods as well as in the Results and the Discussion sections accordingly. I believe these changes are sufficient to visually convey the underlying uncertainty.

### 2019-05-26 article/02_Materials_and_Methods.Rmd
modified figure caption

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
\caption{\label{correlationtimeseries}Development of correlation between different measures of cultural and spatial distance. The results for the correlation of cultural and spatial distance (Table \ref{tab:dacta}) are displayed with a solid (\textit{burial type}) and a dashed (\textit{burial construction}) line. The results for the correlation of the two cultural distances (Table \ref{tab:dactb}) are displayed with a strong dotted line \textit{without} the correction for spatial distance (only here: Pearson correlation coefficient), and with a weak dotted line for the result \textit{with} the aforementioned correction. The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed. Statistically significant correlation (by p-Value) is indicated as such.}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{correlation_time_series.jpeg}
\caption{\label{correlationtimeseries}(\textbf{A}) Development of the median numbers of graves per region and year for the variables \textit{burial type} and \textit{burial construction}. The ribbon indicates the range between the lower and the upper quartile. (\textbf{B}) Development of the correlation between different measures of cultural and spatial distance. The results for the correlation of cultural and spatial distance (Table \ref{tab:dacta}) are displayed with a solid (\textit{burial type}) and a dashed (\textit{burial construction}) line. The results for the correlation of the two cultural distances (Table \ref{tab:dactb}) are displayed with a strong dotted line \textit{without} the correction for spatial distance (only here: Pearson correlation coefficient), and with a weak dotted line for the result \textit{with} the aforementioned correction. The horizontal line marks the correlation coefficient of zero, where neither positive nor negative correlation can be assumed. Statistically significant correlation (by p-Value) is indicated as such. The intensity of the small rectangle around each observation visualizes the mean of the medians in (A) for each 200 years time window and the respective variables.}
\end{figure}
